\documentclass[11pt]{article}
\usepackage{../EllioStyle}

\title{Self/Peer Evaluation}
\author{Elliott Pryor}
\date{18 Nov 2020}


\begin{document}
\maketitle

Overall, I think we did very well as a group. In all honesty, the projects were too small for 5 people so it is kind of unreasonable to expect everyone to contribute a large portion of the project. I think the scope would have been reasonable for a single person, and quite easy with 2-3. That said, I think everyone did a good job contributing and balancing everything. We were a good group.

I did a lot with the group. I helped set up some of the meetings and organized the final products. 
I created a draft of questions that several of our final questions came from. 
I attended every group meeting and helped make sure deliverables were done. I did the screencapture for the recording and made the overleaf document that we worked on. 

I think my groupmates did very well. Jess and Natalie helped a lot with organization and on the questions and general leadership. They responded well on the group Microsoft teams and turned things in on D2L. Arden made an alternate set of questions that many of our final ones were based on. All of them were also at every group call and worked a lot while on the video call. Riley made a set of questions and also initial video for 1.a, 1.b. We didn't end up using the video for continuity. Due to scheduling they weren't at many of the meetings, but still participated through teams. Kyle was fairly absent (mostly in lecture), but he participated in our group calls. He helped make the draft and the recording. 

As I said at the beginning. I think we have a good group and I have absolutely 0 complaints about my groupmates. I would work with all of them again anytime! 

\end{document}


